Figures from: Impacts of host availability and temperature on mosquito-borne parasite transmission

### 2023-12-19 manuscript-figures.Rmd
match up figure numbers with ms

```{r fig-S0_MosquitoDensity, dev = c('pdf','svg', 'png'),  out.width='100%', fig.height=6, fig.width=9, fig.align='center', fig.cap='\\label{fig:MosquitoDensity} Equilibrium mosquito population density, $V^*(T)$, as a function of temperature. Shown here are the median values of $V^*(T)$ with respect to the mosquito trait TPC hyperparameter posterior distributions. Because infection was not assumed to affect life history traits, the \\textit{Aedes aegypti} and DENV and \\textit{Ae. aegypti} and ZIKV curves overlap. As shown in Equation (2.1), $V^*(T)$ is a function of the immature mosquito development rate, probability of immature mosquito survival to adulthood, eggs per female per day, the adult mortality rate, and the carrying capacity for larval mosquitoes. This carrying capacity was assumed to be temperature-independent and set to 300, which resulted in a plausible range of mosquito population densities for the five systems. '}

@@

```{r fig-S1_R0_heatmap_plots_all, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=11, fig.align='center', fig.cap='\\label{fig:R0_heatmap_plots_all} $\\mathcal{R}_0$ as a function of temperature and biting tolerance  ($\\sigma_H$) across three levels of vertebrate host population density: $0.1$, $1$, and $10$ vertebrate hosts per hectare. Shown here are the median values of $\\mathcal{R}_0$ with respect to mosquito trait TPC hyperparameter posterior distributions. Vertebrate host biting tolerance is varied on a $\\log_{10}$ scale with a break to indicate values of $\\mathcal{R}_0$ when biting tolerance is unlimited.'}

@@

```{r fig-S2_Topt_plots_varyKH, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_plots_varyKH} The transmission thermal optimum, $T_{\\textrm{opt}}$, is a sigmoidal function of vertebrate host population density. The variation in $T_{\\textrm{opt}}$ varies substantially between the five systems, with \\textit{Anopheles gambiae} and \\textit{Plasmodium falciparum} (d) having an over 5\\textdegree C change in $T_{\\textrm{opt}}$ as vertebrate host population density is increased. Note that the thermal optimum can be an increasing (d) or decreasing (a, b, c, e) function of vertebrate host population density. Shown here are the median values of $T_{\\textrm{opt}}$ with respect to mosquito trait TPC hyperparameter posterior distributions. Vertebrate host population density is varied on a $\\log_{10}$ scale. All remaining parameters are given in Table 1 and Figure 3.'}

@@

```{r fig-S3_Topt_plots_varysigmaH, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_plots} The thermal optima predicted by the model can be substantially different at low levels of biting tolerance. At low host population densities, there may be warmer (a, b, c, e) or cooler (d) thermal optima when biting tolerance is low compared to when it is high. Shown here are the median values of $T_{\\textrm{opt}}$ with respect to mosquito trait TPC hyperparameter posterior distributions. Vertebrate host biting tolerance is varied on a $\\log_{10}$ scale with a break to indicate values of $\\mathcal{R}_0$ when biting tolerance is unlimited. '}

@@

```{r fig-S4_CTmin, dev = c('pdf','svg', 'png'), dev="cairo_pdf", message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CTmin} The critical thermal minimum for parasite population persistence as a function of vertebrate host population density and biting tolerance. Grey areas indicate combinations of vertebrate host population density and biting tolerance in which the parasite population is expected to not persist at any temperature. In the region where vertebrate host population density and biting tolerance are both less than 10, critical thermal minima and maxima are very similar (see Figure 6 in the main text). Shown here are the median values of the critical thermal minimum with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited. '}

@@

```{r fig-S5_CTmax, dev = c('pdf','svg', 'png'), dev="cairo_pdf", message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CTmax} The critical thermal maximum for parasite population persistence as a function of vertebrate host population density and biting tolerance. Grey areas indicate combinations of vertebrate host population density and biting tolerance in which the parasite population is expected to not persist at any temperature. In the region where vertebrate host population density and biting tolerance are both less than 10, critical thermal minima and maxima are very similar (see Figure 6 in the main text). Shown here are the median values of the critical thermal maximum with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited.'}

@@

```{r fig-S6_CT_Range, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_Range} The critical thermal minimum ($\\textrm{CT}_{\\min}$, blue curves) and maximum ($\\textrm{CT}_{\\max}$, red curves) for parasite population persistence depend on vertebrate host availability, even when vertebrate hosts tolerate any level of mosquito biting (dotted curves). Transmission is sustained ($\\mathcal{R}_0$>1) in the regions enclosed by the red and blue curves. $\\textrm{CT}_{\\min}$ is a convex function of vertebrate host population density while $\\textrm{CT}_{\\max}$ is concave. The points where the red and blue curves touch indicate thresholds of vertebrate host population density beyond which parasite population persistence is no longer possible at any temperature (i.e. $\\mathcal{R}_0$(T)<1) (see Figure 6 in the main text). Shown here are the median values of the critical thermal minimum, maximum, and mosquito thermal tolerance niche with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited. All remaining parameters are given in Table 1 and Figure 2.'}

@@

```{r fig-S7_CT_distributions, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_dists} The distributions of the critical thermal minimum and maximum and thermal optimum vary across transmission systems and levels of vertebrate host availability. Distributions for the limited biting tolerance case (solid lines) and unlimited biting tolerance case (dashed lines) are shown above and below the vertebrate host population density levels, respectively. As vertebrate host population density is increased, the distributions become approximately equal, with the critical thermal minimum and maximum converging to each other at sufficiently high population densities. The \\textit{An. gambiae} and \\textit{P. falciparum} system shows the most drastic differences between the limited and unlimited biting tolerance cases at low vertebrate host population density levels. Vertebrate host population density is varied on a $\\log_{10}$ scale.'}

@@

```{r fig-S8_R0_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:dR0dvar_unc} The local sensitivity of $\\mathcal{R}_0$ to mosquito traits across temperatures for two values of biting tolerance ($\\sigma_H$) and vertebrate host population density ($K_H$). Overall, the sensitivity of $\\mathcal{R}_0$ to temperature decreases at higher temperatures. At high temperatures, adult mortality rate is the primary contributor to the temperature sensitivity of $\\mathcal{R}_0$ across all systems and all values of $\\sigma_H$ and $K_H$. At lower temperatures, the most important trait varies between the adult mortality rate, biting frequency, parasite development rate, or vector competence, depending on the system in question and values of the vertebrate host traits. The values shown here are the median values taken across the mosquito trait TPC hyperparameter posterior distributions of the contributions of each trait to the local sensitivity of derivative of $\\mathcal{R}_0$ to temperature: $\\frac{1}{\\mathcal{R}_0}\\frac{\\partial\\mathcal{R}_0}{\\partial p}\\frac{dp}{dT}$ where $p$ is a focal trait and $T$ is temperature. Values were scaled to span from -1 to 1 to enable comparisons across systems. Temperatures are restricted to those at which $\\mathcal{R}_0$ is positive for the given system. '}

>>>>

```{r fig-S1_MosquitoDensity, dev = c('pdf','svg', 'png'),  out.width='100%', fig.height=6, fig.width=9, fig.align='center', fig.cap='\\label{fig:MosquitoDensity} Equilibrium mosquito population density, $V^*(T)$, as a function of temperature. Shown here are the median values of $V^*(T)$ with respect to the mosquito trait TPC hyperparameter posterior distributions. Because infection was not assumed to affect life history traits, the \\textit{Aedes aegypti} and DENV and \\textit{Ae. aegypti} and ZIKV curves overlap. As shown in Equation (2.1), $V^*(T)$ is a function of the immature mosquito development rate, probability of immature mosquito survival to adulthood, eggs per female per day, the adult mortality rate, and the carrying capacity for larval mosquitoes. This carrying capacity was assumed to be temperature-independent and set to 300, which resulted in a plausible range of mosquito population densities for the five systems. '}

@@

```{r fig-S2_R0_heatmap_plots_all, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=11, fig.align='center', fig.cap='\\label{fig:R0_heatmap_plots_all} $\\mathcal{R}_0$ as a function of temperature and biting tolerance  ($\\sigma_H$) across three levels of vertebrate host population density: $0.1$, $1$, and $10$ vertebrate hosts per hectare. Shown here are the median values of $\\mathcal{R}_0$ with respect to mosquito trait TPC hyperparameter posterior distributions. Vertebrate host biting tolerance is varied on a $\\log_{10}$ scale with a break to indicate values of $\\mathcal{R}_0$ when biting tolerance is unlimited.'}

@@

```{r fig-S3_Topt_plots_varyKH, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_plots_varyKH} The transmission thermal optimum, $T_{\\textrm{opt}}$, is a sigmoidal function of vertebrate host population density. The variation in $T_{\\textrm{opt}}$ varies substantially between the five systems, with \\textit{Anopheles gambiae} and \\textit{Plasmodium falciparum} (d) having an over 5\\textdegree C change in $T_{\\textrm{opt}}$ as vertebrate host population density is increased. Note that the thermal optimum can be an increasing (d) or decreasing (a, b, c, e) function of vertebrate host population density. Shown here are the median values of $T_{\\textrm{opt}}$ with respect to mosquito trait TPC hyperparameter posterior distributions. Vertebrate host population density is varied on a $\\log_{10}$ scale. All remaining parameters are given in Table 1 and Figure 3.'}

@@

```{r fig-S4_Topt_plots_varysigmaH, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_plots} The thermal optima predicted by the model can be substantially different at low levels of biting tolerance. At low host population densities, there may be warmer (a, b, c, e) or cooler (d) thermal optima when biting tolerance is low compared to when it is high. Shown here are the median values of $T_{\\textrm{opt}}$ with respect to mosquito trait TPC hyperparameter posterior distributions. Vertebrate host biting tolerance is varied on a $\\log_{10}$ scale with a break to indicate values of $\\mathcal{R}_0$ when biting tolerance is unlimited. '}

@@

```{r fig-S5_CTmin, dev = c('pdf','svg', 'png'), dev="cairo_pdf", message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CTmin} The critical thermal minimum for parasite population persistence as a function of vertebrate host population density and biting tolerance. Grey areas indicate combinations of vertebrate host population density and biting tolerance in which the parasite population is expected to not persist at any temperature. In the region where vertebrate host population density and biting tolerance are both less than 10, critical thermal minima and maxima are very similar (see Figure 6 in the main text). Shown here are the median values of the critical thermal minimum with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited. '}

@@

```{r fig-S6_CTmax, dev = c('pdf','svg', 'png'), dev="cairo_pdf", message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CTmax} The critical thermal maximum for parasite population persistence as a function of vertebrate host population density and biting tolerance. Grey areas indicate combinations of vertebrate host population density and biting tolerance in which the parasite population is expected to not persist at any temperature. In the region where vertebrate host population density and biting tolerance are both less than 10, critical thermal minima and maxima are very similar (see Figure 6 in the main text). Shown here are the median values of the critical thermal maximum with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited.'}

@@

```{r fig-S7_CT_Range, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_Range} The critical thermal minimum ($\\textrm{CT}_{\\min}$, blue curves) and maximum ($\\textrm{CT}_{\\max}$, red curves) for parasite population persistence depend on vertebrate host availability, even when vertebrate hosts tolerate any level of mosquito biting (dotted curves). Transmission is sustained ($\\mathcal{R}_0$>1) in the regions enclosed by the red and blue curves. $\\textrm{CT}_{\\min}$ is a convex function of vertebrate host population density while $\\textrm{CT}_{\\max}$ is concave. The points where the red and blue curves touch indicate thresholds of vertebrate host population density beyond which parasite population persistence is no longer possible at any temperature (i.e. $\\mathcal{R}_0$(T)<1) (see Figure 6 in the main text). Shown here are the median values of the critical thermal minimum, maximum, and mosquito thermal tolerance niche with respect to mosquito trait TPC hyperparameter posterior distributions. Both vertebrate host population density and biting tolerance are varied on a $\\log_{10}$ scale with a break in biting tolerance values to indicate when biting tolerance is unlimited. All remaining parameters are given in Table 1 and Figure 2.'}

@@

```{r fig-S8_CT_distributions, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_dists} The distributions of the critical thermal minimum and maximum and thermal optimum vary across transmission systems and levels of vertebrate host availability. Distributions for the limited biting tolerance case (solid lines) and unlimited biting tolerance case (dashed lines) are shown above and below the vertebrate host population density levels, respectively. As vertebrate host population density is increased, the distributions become approximately equal, with the critical thermal minimum and maximum converging to each other at sufficiently high population densities. The \\textit{An. gambiae} and \\textit{P. falciparum} system shows the most drastic differences between the limited and unlimited biting tolerance cases at low vertebrate host population density levels. Vertebrate host population density is varied on a $\\log_{10}$ scale.'}

@@

```{r fig-S9_R0_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:dR0dvar_unc} The local sensitivity of $\\mathcal{R}_0$ to mosquito traits across temperatures for two values of biting tolerance ($\\sigma_H$) and vertebrate host population density ($K_H$). Overall, the sensitivity of $\\mathcal{R}_0$ to temperature decreases at higher temperatures. At high temperatures, adult mortality rate is the primary contributor to the temperature sensitivity of $\\mathcal{R}_0$ across all systems and all values of $\\sigma_H$ and $K_H$. At lower temperatures, the most important trait varies between the adult mortality rate, biting frequency, parasite development rate, or vector competence, depending on the system in question and values of the vertebrate host traits. The values shown here are the median values taken across the mosquito trait TPC hyperparameter posterior distributions of the contributions of each trait to the local sensitivity of derivative of $\\mathcal{R}_0$ to temperature: $\\frac{1}{\\mathcal{R}_0}\\frac{\\partial\\mathcal{R}_0}{\\partial p}\\frac{dp}{dT}$ where $p$ is a focal trait and $T$ is temperature. Values were scaled to span from -1 to 1 to enable comparisons across systems. Temperatures are restricted to those at which $\\mathcal{R}_0$ is positive for the given system. '}
